---
title: "Data Analysis and Visualization Project"
author: "Jie HAN"
date: "11/29/2018"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

1. Set work direction

```{r}
setwd <- ("/Users/jill/Downloads/631_R/assignment/project/")
library(ggplot2)
library(stats)
library (dplyr)
library(corrplot)
library (dplyr)
library(psych)
library(rgl)
library(pwr)
library(pscl)
library(ISLR)
library(dlookr)
library(Hmisc)
library(pastecs)
library(car)
logisticPseudoR2s <- function(LogModel) {
  dev <- LogModel$deviance 
  nullDev <- LogModel$null.deviance 
  modelN <-  length(LogModel$fitted.values)
  R.l <-  1 -  dev / nullDev
  R.cs <- 1- exp ( -(nullDev - dev) / modelN)
  R.n <- R.cs / ( 1 - ( exp (-(nullDev / modelN))))
  cat("Pseudo R^2 for logistic regression\n")
  cat("Hosmer and Lemeshow R^2  ", round(R.l, 3), "\n")
  cat("Cox and Snell R^2        ", round(R.cs, 3), "\n")
  cat("Nagelkerke R^2           ", round(R.n, 3),    "\n")
}
```


2. Importing the data set

```{r}
diab0 <- read.csv("diabetic_data.csv", header = T, na.strings = "?")
class(diab0)
names(diab0)
dim(diab0)
str(diab0)
summary(diab0)
```


3. Cleaning data 

```{r}
# Remove the repeate
diab <- diab0 %>% group_by(patient_nbr) %>% filter(encounter_id == min(encounter_id)) 

# Remove "weight" variables because of too many unreliable values.
# Remove meaningless variables ("encounter_id","patient_nbr")
# Remove useless or repetitive meaning variables( "max_glu_serum", 25:41 43:47 variables are medicines)

diab2 <- diab[-c(1:2,6,23,25:41,43:47)]
summary(diab2)
names(diab2)
```


## Check the ordered factor levels
```{r}
# age
levels(diab2$age)
# "[0-10)"   "[10-20)"  "[20-30)"  "[30-40)"  "[40-50)"  "[50-60)"  "[60-70)"  "[70-80)"  "[80-90)"  "[90-100)" 
# The order is correct, but too many levels, collapse them into three levels: [0-30), [30~70),[70~100)
ind_30 <- which(diab2$age == "[0-10)"|diab2$age ==  "[10-20)"| diab2$age =="[20-30)")
ind_70 <- which(diab2$age == "[70-80)"|diab2$age ==  "[80-90)"| diab2$age =="[90-100)")

diab2$new_age <- factor (rep(NA, nrow(diab2)), order=TRUE, levels= c("[0-30)", "[30~70)","[70~100)"))
diab2$new_age[ind_30]<- "[0-30)"
diab2$new_age[ind_70]<- "[70~100)"
diab2$new_age[-c(ind_30,ind_70)] <- "[30~70)"
table(diab2$new_age)
```


```{r}
# A1Cresult
levels(diab2$A1Cresult)
# ">7"   ">8"   "None" "Norm". The order is not correct, need to replace its levels.
A1C_lev <- c("None", "Norm", ">7",">8")
diab2$A1Cresult <- factor(diab2$A1Cresult, levels=A1C_lev, order= TRUE)
levels(diab2$A1Cresult)
```


```{r}
# Insulin 
diab2$insulin <- ordered(diab2$insulin)
# "Down"   "No"     "Steady" "Up". The order level is correct.
```

```{r}
# readmitted
levels(diab2$readmitted)
# "<30" ">30" "NO". The order level should be changed to "no","<30", ">30"
readmit_lev <- c("NO","<30", ">30")
diab2$readmitted <- factor(diab2$readmitted, levels = readmit_lev,order= TRUE)
levels(diab2$readmitted)
```
## The wrong type of variables: "admission_type_id","discharge_disposition_id","admission_source_id" should be factor
```{r}
diab2$admission_type_id <- factor(diab2$admission_type_id)
diab2$discharge_disposition_id <- factor(diab2$discharge_disposition_id)
diab2$admission_source_id <- factor(diab2$admission_source_id)
```



## Collapse levels: "gender", "admission_type_id","discharge_disposition_id","admission_source_id","payer_code","medical_specialty","diag_1","diag_2","diag_3"

```{r}
levels(diab2$gender)
diab2$gender[which(diab2$gender == "Unknown/Invalid")] <- NA
gen_lev <- c("Female","Male")
diab2$gender <- factor(diab2$gender, levels = gen_lev)
summary(diab2$gender)
```

```{r}
# admission_type_id
lev1 <- c("Emergency","Urgent","Elective","Newborn","Not Available","NULL","Trauma Center","Not Mapped")

diab2$new_adtype_id <- factor(diab2$admission_type_id, labels = lev1)
summary(diab2$new_adtype_id)

no_keep_ind <- which(diab2$new_adtype_id=="Not Available" | diab2$new_adtype_id=="NULL" |
                  diab2$new_adtype_id=="Not Mapped")

diab2$new_adtype_id[no_keep_ind] <- "Not Mapped"
diab2$new_adtype_id <- factor(diab2$new_adtype_id, levels =c("Emergency","Urgent","Elective","Newborn","Trauma Center","Not Mapped"))

summary(diab2$new_adtype_id)
```

```{r}
# discharge_disposition_id
summary(diab2$discharge_disposition_id)
diab2$new_discharge_id <-  factor (rep(NA, nrow(diab2)), levels= c("hospice", "expired","others"))

hos_ind <- which(diab2$discharge_disposition_id == "13"|diab2$discharge_disposition_id == "14")
exp_ind <- which(diab2$discharge_disposition_id == "11"|diab2$discharge_disposition_id == "19"|
                 diab2$discharge_disposition_id == "20"|diab2$discharge_disposition_id == "21" )

diab2$new_discharge_id[hos_ind] <- "hospice"
diab2$new_discharge_id[exp_ind] <- "expired"
diab2$new_discharge_id[-c(hos_ind,exp_ind)] <- "others"

summary(diab2$new_discharge_id)
```

```{r}
# admission_source_id
summary(diab2$admission_source_id)
diab2$new_adsource_id <-  factor (rep(NA, nrow(diab2)), levels= c("referral", "emergency", "transfer","others"))

ref_ind <- which(diab2$admission_source_id == "1"|diab2$admission_source_id == "2"|diab2$admission_source_id == "3")
emer_ind <- which(diab2$admission_source_id == "7")
tran_ind <- which(diab2$admission_source_id == "4"|diab2$admission_source_id == "5"|diab2$admission_source_id == "6"|
                    diab2$admission_source_id == "10"|diab2$admission_source_id == "18"|diab2$admission_source_id == "22"|
                    diab2$admission_source_id == "25"|diab2$admission_source_id == "26")

diab2$new_adsource_id[ref_ind] <- "referral"
diab2$new_adsource_id[emer_ind] <- "emergency"
diab2$new_adsource_id[tran_ind] <- "transfer"
diab2$new_adsource_id[-c(ref_ind,emer_ind,tran_ind)] <- "others"

summary(diab2$new_adsource_id)
```

```{r}
# medical_specialty

summary(diab2$medical_specialty)  # Check if there are NAs 
diab2$new_med_spec <-  factor (rep(NA, nrow(diab2)), levels=c("Cardiology","Emergency/Trauma","InternalMedicine","others"))

car_ind <- which(diab2$medical_specialty == "Cardiology")
eme_ind <- which(diab2$medical_specialty =="Emergency/Trauma")
inter_ind <- which( diab2$medical_specialty == "InternalMedicine")
na_ind <- which(is.na(diab2$medical_specialty))

diab2$new_med_spec[car_ind] <- "Cardiology"
diab2$new_med_spec[eme_ind] <- "Emergency/Trauma"    
diab2$new_med_spec[inter_ind] <- "InternalMedicine" 
diab2$new_med_spec[-c(car_ind,eme_ind, inter_ind,na_ind)] <- "others"

summary(diab2$new_med_spec)
```


## ICD9 codes:
   Circulatory disease (icd9 : 390–459, 785)
   Diabetes (icd9 : 250.xx)
   Respiratory disease (icd9 : 460–519, 786)
   Digestive disease (icd9 : 520–579, 787)
   Injury, poisoning (icd9 : 800–999)
   Genital or urinary diseases (icd9 : 580–629, 788)
   Neoplasms (icd9:140-239,780,781,784,790-799,240-249,251-279)
   

```{r}
# diag_1

summary(diab2$diag_1)
diab2$diag_1d <- as.numeric(as.character(diab2$diag_1))
summary(diab2$diag_1d)
diab2$new_diag1 <- factor (rep(NA, nrow(diab2)), levels=c("Cir","Dia","Res","Dig","Inj","Gen","Can","Others"))

cir_ind1 <- c(which(diab2$diag_1d >= 390 & diab2$diag_1d <= 459), which(diab2$diag_1d == 785))
dia_ind1 <- which(diab2$diag_1d >= 250 & diab2$diag_1d < 251)
res_ind1 <- c(which(diab2$diag_1d >= 460 & diab2$diag_1d <= 519),which(diab2$diag_1d == 786))
dig_ind1 <- c(which(diab2$diag_1d >= 520 & diab2$diag_1d <= 579),which(diab2$diag_1d == 787))
inj_ind1 <- which(diab2$diag_1d >= 800 & diab2$diag_1d <= 999 )
gen_ind1 <- c(which(diab2$diag_1d >= 580 & diab2$diag_1d <= 629),which(diab2$diag_1d == 788 ))
can_ind1 <- c(which(diab2$diag_1d >= 140 & diab2$diag_1d <= 239),which(diab2$diag_1d >= 240 & diab2$diag_1d <= 249),
             which(diab2$diag_1d >= 251 & diab2$diag_1d <= 279),which(diab2$diag_1d >= 790 & diab2$diag_1d <= 799),
             which(diab2$diag_1d == 780| diab2$diag_1d == 781|diab2$diag_1d == 784))
na_ind1 <- which(is.na(diab2$diag_1d))

diab2$new_diag1[cir_ind1] <- "Cir"
diab2$new_diag1[dia_ind1] <- "Dia"
diab2$new_diag1[res_ind1] <- "Res"
diab2$new_diag1[dig_ind1] <- "Dig"
diab2$new_diag1[inj_ind1] <- "Inj"
diab2$new_diag1[gen_ind1] <- "Gen"
diab2$new_diag1[can_ind1] <- "Can"
diab2$new_diag1[-c(cir_ind1,dia_ind1,res_ind1,dig_ind1,inj_ind1,gen_ind1,can_ind1,na_ind1)] <- "Others"

summary(diab2$new_diag1)

```

```{r}
# diag_2

summary(diab2$diag_2)
diab2$diag_2d <- as.numeric(as.character(diab2$diag_2))
diab2$new_diag2 <- factor (rep(NA, nrow(diab2)), levels=c("Cir","Dia","Res","Dig","Inj","Gen","Can","Others"))

cir_ind2 <- c(which(diab2$diag_2d >= 390 & diab2$diag_2d <= 459), which(diab2$diag_2d == 785))
dia_ind2 <- which(diab2$diag_2d >= 250 & diab2$diag_2d < 251)
res_ind2 <- c(which(diab2$diag_2d >= 460 & diab2$diag_2d <= 519),which(diab2$diag_2d == 786))
dig_ind2 <- c(which(diab2$diag_2d >= 520 & diab2$diag_2d <= 579),which(diab2$diag_2d == 787))
inj_ind2 <- which(diab2$diag_2d >= 800 & diab2$diag_2d <= 999 )
gen_ind2 <- c(which(diab2$diag_2d >= 580 & diab2$diag_2d <= 629),which(diab2$diag_2d == 788 ))
can_ind2 <- c(which(diab2$diag_2d >= 140 & diab2$diag_2d <= 239),which(diab2$diag_2d >= 240 & diab2$diag_2d <= 249),
             which(diab2$diag_2d >= 251 & diab2$diag_2d <= 279),which(diab2$diag_2d >= 790 & diab2$diag_2d <= 799),
             which(diab2$diag_2d == 780| diab2$diag_2d == 781|diab2$diag_2d == 784))
na_ind2 <- which(is.na(diab2$diag_2d))

diab2$new_diag2[cir_ind2] <- "Cir"
diab2$new_diag2[dia_ind2] <- "Dia"
diab2$new_diag2[res_ind2] <- "Res"
diab2$new_diag2[dig_ind2] <- "Dig"
diab2$new_diag2[inj_ind2] <- "Inj"
diab2$new_diag2[gen_ind2] <- "Gen"
diab2$new_diag2[can_ind2] <- "Can"
diab2$new_diag2[-c(cir_ind2,dia_ind2,res_ind2,dig_ind2,inj_ind2,gen_ind2,can_ind2,na_ind2)] <- "Others"

summary(diab2$new_diag2)

```

```{r}
# diag_3
summary(diab2$diag_3)
diab2$diag_3d <- as.numeric(as.character(diab2$diag_3))
summary(diab2$diag_3d)
diab2$new_diag3 <- factor (rep(NA, nrow(diab2)), levels=c("Cir","Dia","Res","Dig","Inj","Gen","Can","Others"))

cir_ind3 <- c(which(diab2$diag_3d >= 390 & diab2$diag_3d <= 459), which(diab2$diag_3d == 785))
dia_ind3 <- which(diab2$diag_3d >= 250 & diab2$diag_3d < 251)
res_ind3 <- c(which(diab2$diag_3d >= 460 & diab2$diag_3d <= 519),which(diab2$diag_3d == 786))
dig_ind3 <- c(which(diab2$diag_3d >= 520 & diab2$diag_3d <= 579),which(diab2$diag_3d == 787))
inj_ind3 <- which(diab2$diag_3d >= 800 & diab2$diag_3d <= 999 )
gen_ind3 <- c(which(diab2$diag_3d >= 580 & diab2$diag_3d <= 629),which(diab2$diag_3d == 788 ))
can_ind3 <- c(which(diab2$diag_3d >= 140 & diab2$diag_3d <= 239),which(diab2$diag_3d >= 240 & diab2$diag_3d <= 249),
             which(diab2$diag_3d >= 251 & diab2$diag_3d <= 279),which(diab2$diag_3d >= 790 & diab2$diag_3d <= 799),
             which(diab2$diag_3d == 780| diab2$diag_3d == 781|diab2$diag_3d == 784))
na_ind3 <- which(is.na(diab2$diag_3d))

diab2$new_diag3[cir_ind3] <- "Cir"
diab2$new_diag3[dia_ind3] <- "Dia"
diab2$new_diag3[res_ind3] <- "Res"
diab2$new_diag3[dig_ind3] <- "Dig"
diab2$new_diag3[inj_ind3] <- "Inj"
diab2$new_diag3[gen_ind3] <- "Gen"
diab2$new_diag3[can_ind3] <- "Can"
diab2$new_diag3[-c(cir_ind3,dia_ind3,res_ind3,dig_ind3,inj_ind3,gen_ind3,can_ind3,na_ind3)] <- "Others"

summary(diab2$new_diag3)

```



## Final dataframe to test
```{r}
diab_test <- diab2[-c(3:6,8:9,16:18,30,32,34)]
summary(diab_test)
apply(diab_test[3:10],2,var,na.rm = TRUE)
apply(diab_test[3:10],2,sd, na.rm=TRUE)
describe(diab_test)
```


##  Univariate data EDA

4. Visualization

a. Distributions of each numerical variable 

```{r}
g <- ggplot(diab_test)
g + geom_bar(aes(x=race),fill = 1,color="grey") +
  ggtitle("Distribution of Race")+xlab("Race")+ ylab("Count")

g + geom_bar(aes(x=gender),fill = 1,color="grey") +
  ggtitle("Distribution of Gender")+xlab("Gender")+ ylab("Count")

g + geom_histogram(aes(x=time_in_hospital), binwidth = 1,fill = 1,color="grey")+
  ggtitle("Distribution of Time_in_hospital")+xlab("Time_in_hospital (days)")+ ylab("Frequency")

g + geom_histogram(aes(x=num_lab_procedures), binwidth = 10,fill = 1,color="grey")+
  ggtitle("Distribution of Num_lab_procedures")+xlab("Num_lab_procedures")+ ylab("Frequency")

g + geom_histogram(aes(x=num_procedures), binwidth = 1,fill = 1,color="grey") +
  ggtitle("Distribution of Num_procedures")+xlab("Num_procedures")+ ylab("Frequency")

g + geom_histogram(aes(x=num_medications),binwidth = 10,fill = 1,color="grey") +
  ggtitle("Distribution of Num_medications")+xlab("Num_medications")+ ylab("Frequency")

g + geom_histogram(aes(x=number_outpatient),binwidth = 10,fill = 1,color="grey") +
  ggtitle("Distribution of Number_outpatient")+xlab("Number_outpatient")+ ylab("Frequency")

g + geom_histogram(aes(x=number_emergency),binwidth = 2,fill = 1,color="grey") +
  ggtitle("Distribution of Number_emergency")+xlab("Number_emergency")+ ylab("Frequency")

g + geom_histogram(aes(x=number_inpatient),binwidth = 2,fill = 1,color="grey") +
  ggtitle("Distribution of Number_inpatient")+xlab("Number_inpatient")+ ylab("Frequency")

g + geom_histogram(aes(x=number_diagnoses),binwidth = 2,fill = 1,color="grey") +
  ggtitle("Distribution of Number_diagnoses")+xlab("Number_diagnoses")+ ylab("Frequency")

g + geom_bar(aes(x=A1Cresult),fill = 1,color="grey") +
  ggtitle("Distribution of A1Cresult")+xlab("A1Cresult")+ ylab("Count")

g + geom_bar(aes(x=insulin),fill = 1,color="grey") +
  ggtitle("Distribution of Insulin")+xlab("Insulin")+ ylab("Count")

g + geom_bar(aes(x=change),fill = 1,color="grey") +
  ggtitle("Distribution of Change")+xlab("Change")+ ylab("Count")

g + geom_bar(aes(x=diabetesMed),fill = 1,color="grey") +
  ggtitle("Distribution of DiabetesMed")+xlab("DiabetesMed")+ ylab("Count")

g + geom_bar(aes(x=readmitted),fill = 1,color="grey") +
  ggtitle("Distribution of Readmitted")+xlab("Readmitted")+ ylab("Count")

g + geom_bar(aes(x=new_age),fill = 1,color="grey") +
  ggtitle("Distribution of New_age")+xlab("New_age")+ ylab("Count")

g + geom_bar(aes(x=new_adtype_id),fill = 1,color="grey") +
  ggtitle("Distribution of New_adtype_id")+xlab("New_adtype_id")+ ylab("Count")

g + geom_bar(aes(x=new_discharge_id),fill = 1,color="grey") +
  ggtitle("Distribution of New_discharge_id")+xlab("New_discharge_id")+ ylab("Count")

g + geom_bar(aes(x=new_adsource_id),fill = 1,color="grey") +
  ggtitle("Distribution of New_adsource_id")+xlab("New_adsource_idt")+ ylab("Count")

g + geom_bar(aes(x=new_med_spec),fill = 1,color="grey") +
  ggtitle("Distribution of New_med_spe")+xlab("New_med_spe")+ ylab("Count")

g + geom_bar(aes(x=new_diag1),fill = 1,color="grey") +
  ggtitle("Distribution of New_diag1")+xlab("New_diag1")+ ylab("Count")

g + geom_bar(aes(x=new_diag2),fill = 1,color="grey") +
  ggtitle("Distribution of New_diag2")+xlab("New_diag1")+ ylab("Count")

g + geom_bar(aes(x=new_diag3),fill = 1,color="grey") +
  ggtitle("Distribution of New_diag3")+xlab("New_diag1")+ ylab("Count")
```


b. Distributions of each numerical variable with fill color determined by readmitted
```{r}
a <- Filter(is.numeric, diab_test)
names(a)
# "time_in_hospital"   "num_lab_procedures" "num_procedures"     "num_medications"    "number_outpatient" 
# "number_emergency"   "number_inpatient"   "number_diagnoses"
g <- ggplot(diab_test)
# should plot each variable with 2 plots?
g + geom_boxplot(aes(x=readmitted, y=time_in_hospital,fill=readmitted))
g + geom_histogram(aes(x=time_in_hospital,fill=readmitted), binwidth = 1,position = "dodge")+
  ggtitle("Distribution of Time_in_hospital")+
  xlab("time_in_hospital (days)")+ ylab("Frequency")

g + geom_boxplot(aes(x=readmitted, y=num_lab_procedures,fill=readmitted))
g + geom_histogram(aes(x=num_lab_procedures,fill=readmitted), binwidth = 10,position = "dodge")+
  ggtitle("Distribution of Num_lab_procedures")+
  xlab("Num_lab_procedures")+ ylab("Frequency")

g + geom_boxplot(aes(x=readmitted, y=num_procedures,fill=readmitted))
g + geom_histogram(aes(x=num_procedures,fill=readmitted), binwidth = 1,position = "dodge")+
  ggtitle("Distribution of Num_procedures")+
  xlab("Num_procedures")+ ylab("Frequency")


g + geom_boxplot(aes(x=readmitted, y=num_medications,fill=readmitted))
g + geom_histogram(aes(x=num_medications,fill=readmitted), binwidth = 10,position = "dodge")+
  ggtitle("Distribution of Number_medications")+
  xlab("Number_medications")+ ylab("Frequency")

g + geom_boxplot(aes(x=readmitted, y=number_outpatient,fill=readmitted))
g + geom_histogram(aes(x=number_outpatient, fill=readmitted), binwidth = 10,position = "dodge")+
  ggtitle("Distribution of Number_outpatient")+
  xlab("Number_outpatient")+ ylab("Frequency")

g + geom_boxplot(aes(x=readmitted, y=number_emergency,fill=readmitted))
g + geom_histogram(aes(x=number_emergency, fill=readmitted), binwidth = 10,position = "dodge")+
  ggtitle("Distribution of Number_emergency")+
  xlab("Number_emergency")+ ylab("Frequency")

g + geom_boxplot(aes(x=readmitted, y=number_inpatient,fill=readmitted))
g + geom_histogram(aes(x=number_inpatient, fill=readmitted), binwidth = 10,position = "dodge")+
  ggtitle("Distribution of Number_inpatient")+
  xlab("Number_inpatient")+ ylab("Frequency")

g + geom_boxplot(aes(x=readmitted, y=number_diagnoses,fill=readmitted))
g + geom_histogram(aes(x=number_diagnoses,fill=readmitted), binwidth = 5,position = "dodge")+
  ggtitle("Distribution of Number_diagnoses")+
  xlab("Number_diagnoses")+ ylab("Frequency")
```


c. Grouped barplots of categorical variables

```{r}
df <- diab_test[,sapply(diab_test, is.factor)]
names(df)
# "race"             "gender"           "payer_code"       "A1Cresult"        "insulin"         
# "change"           "diabetesMed"      "readmitted"       "new_age"          "new_adtype_id"   
# "new_discharge_id" "new_adsource_id"  "new_med_spec"     "new_diag1"        "new_diag2"       
# "new_diag3"  

g <- ggplot(diab_test)

g + geom_bar(aes(x=race, fill= readmitted))+ 
  ggtitle("Barplot of Race")+
  xlab("Race") + ylab("Count")

g + geom_bar(aes(x=gender, fill= readmitted))+
  ggtitle("Barplot of Gender")+
  xlab("Gender") + ylab("Count")

g + geom_bar(aes(x=new_age, fill= readmitted))+
  ggtitle("Barplot of Age")+
  xlab("Age") + ylab("Count")

g + geom_bar(aes(x=new_adtype_id, fill= readmitted))+
  ggtitle("Barplot of New_adtype_id")+
  xlab("New_adtype_id") + ylab("Count")

g + geom_bar(aes(x=new_discharge_id, fill= readmitted))+
  ggtitle("Barplot of New_discharge_id")+
  xlab("New_discharge_id") + ylab("Count")

g + geom_bar(aes(x=new_adsource_id, fill= readmitted))+
  ggtitle("Barplot of New_adsource_id")+
  xlab("New_adsource_id") + ylab("Count")

g + geom_bar(aes(x=new_med_spec, fill= readmitted))+
  ggtitle("Barplot of New_med_spec")+
  xlab("New_med_spec") + ylab("Count")

g + geom_bar(aes(x=new_diag1, fill= readmitted))+
  ggtitle("Barplot of New_diag1")+
  xlab("New_diag1") + ylab("Count")

g + geom_bar(aes(x=new_diag2, fill= readmitted))+
  ggtitle("Barplot of New_diag2")+
  xlab("New_diag2") + ylab("Count")

g + geom_bar(aes(x=new_diag3, fill= readmitted))+
  ggtitle("Barplot of New_diag3")+
  xlab("New_diag3") + ylab("Count")
                           
g + geom_bar(aes(x=A1Cresult, fill= readmitted))+
  ggtitle("Barplot of A1Cresult")+
  xlab("A1Cresult") + ylab("Count")

g + geom_bar(aes(x=insulin, fill= readmitted))+
  ggtitle("Barplot of Insulin")+
  xlab("Insulin") + ylab("Count")

g + geom_bar(aes(x=change, fill= readmitted))+
  ggtitle("Barplot of Change")+
  xlab("Change") + ylab("Count")

g + geom_bar(aes(x=diabetesMed, fill= readmitted))+
  ggtitle("Barplot of DiabetesMed")+
  xlab("DiabetesMed") + ylab("Count")

g + geom_bar(aes(x=readmitted,fill= readmitted))+
  ggtitle("Barplot of Readmitted")+
  xlab("Readmitted") + ylab("Count")

```


5. Summary statistcs

The types of variables:

Continous: time_in_hospital, num_lab_procedures, num_procedures, num_medications, number_outpatient, number_emergency, number_inpatient, number_diagnoses

Binary: gender, change, diabetesMed
Nominal: race, new_diag1, new_diag2, new_diag3, new_adtype_id, new_discharge_id, new_adsource_id, new_med_spec
Ordinal: new_age, A1Cresult, insulin, readmitted

```{r}
summary(diab_test)
describe(diab_test)
```


## Bivariate data EDA

7. Correlations
"time_in_hospital" have correlation with num_lab_procedures, num_procedures,num_medications,num_diagnoses, num_procedures, new_age. 

num_meditions has relationship with num_lab_procedures,num_diagnoses,num_procedures.

```{r}
str(diab_test)
diab_numeric <- diab_test[,3:10]

diab_numeric$new_age <- as.numeric(diab_test$new_age)
diab_numeric$A1Cresult <- as.numeric(diab_test$A1Cresult)
diab_numeric$insulin <- as.numeric(diab_test$insulin)
diab_numeric$readmitted <- as.numeric(diab_test$readmitted)

c <- cor(diab_numeric, method = "spearman", use = "pairwise.complete.obs")
c
corrplot(c,method="circle")
```


7. Correlations test

medium association with time_in_hospital(0.3~0.49): num_lab_procedures, num_medications
small association with time_in_hospital(0.1~0.29): num_diagnoses, num_procedures, new_age

association with readmitted: 

```{r}
# For two numerical variables, use pearson method.
cor.test(diab_test$ num_lab_procedures, diab_test$time_in_hospital, use= "pairwise.complete.obs")
cor.test(diab_test$num_procedures, diab_test$time_in_hospital, use= "pairwise.complete.obs")
cor.test(diab_test$num_medications, diab_test$time_in_hospital, use= "pairwise.complete.obs")
cor.test(diab_test$number_outpatient, diab_test$time_in_hospital, use= "pairwise.complete.obs")
cor.test(diab_test$number_emergency, diab_test$time_in_hospital, use= "pairwise.complete.obs")
cor.test(diab_test$number_inpatient, diab_test$time_in_hospital, use= "pairwise.complete.obs")
cor.test(diab_test$number_diagnoses, diab_test$time_in_hospital, use= "pairwise.complete.obs")

# one numerical and one ordinal variables (spearman for ordinal)
cor.test(as.numeric(diab_test$new_age), diab_test$time_in_hospital,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$A1Cresult), diab_test$time_in_hospital,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$insulin), diab_test$time_in_hospital,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$time_in_hospital,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$num_lab_procedures,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$num_procedures,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$num_medications,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$number_diagnoses,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$number_outpatient,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$number_emergency,use= "pairwise.complete.obs",method="spearman")
cor.test(as.numeric(diab_test$readmitted), diab_test$number_diagnoses,use= "pairwise.complete.obs",method="spearman")

# Two ordinal variables 
cor.test(as.numeric(diab_test$readmitted), as.numeric(diab_test$new_age),use= "pairwise.complete.obs",method="kendall")
cor.test(as.numeric(diab_test$readmitted), as.numeric(diab_test$A1Cresult),use= "pairwise.complete.obs",method="kendall")
cor.test(as.numeric(diab_test$readmitted), as.numeric(diab_test$insulin),use= "pairwise.complete.obs",method="kendall")
```


8. Statiscal tests
HO: no difference between the two true means.

The types of variables:

Continous: time_in_hospital, num_lab_procedures, num_procedures, num_medications, number_outpatient, number_emergency, number_inpatient, number_diagnoses

Binary: gender, change, diabetesMed
Nominal: race, new_diag1, new_diag2, new_diag3, new_adtype_id, new_discharge_id, new_adsource_id, new_med_spec
Ordinal: new_age, A1Cresult, insulin, readmitted


Dependent variables: time_in_hospital
```{r}

t.test(time_in_hospital~gender,data=diab_test)
t.test(time_in_hospital~change,data=diab_test)
t.test(time_in_hospital~diabetesMed,data=diab_test)

oneway.test(time_in_hospital~race,data=diab_test,var.equal=FALSE)
oneway.test(time_in_hospital~new_adtype_id,data=diab_test,var.equal=FALSE)
oneway.test(time_in_hospital~new_discharge_id,data=diab_test,var.equal=FALSE)
oneway.test(time_in_hospital~new_med_spec,data=diab_test,var.equal=FALSE)
oneway.test(time_in_hospital~new_diag1,data=diab_test,var.equal=FALSE)
oneway.test(time_in_hospital~new_diag2,data=diab_test,var.equal=FALSE)
oneway.test(time_in_hospital~new_diag3,data=diab_test,var.equal=FALSE)
```


Denpendent variables: readmitted
```{r}
chisq.test(table(diab_test$readmitted, diab_test$race))
chisq.test(table(diab_test$readmitted, diab_test$new_adtype_id))
chisq.test(table(diab_test$readmitted, diab_test$new_discharge_id))
chisq.test(table(diab_test$readmitted, diab_test$new_med_spec))
chisq.test(table(diab_test$readmitted, diab_test$new_diag1))
chisq.test(table(diab_test$readmitted, diab_test$new_diag2))
chisq.test(table(diab_test$readmitted, diab_test$new_diag3))
```


9. Linear regression model of "time_in_hospital"

As num_meditions has relationship with num_lab_procedures,num_diagnoses,num_procedures, keep num_meditions in regression model 

```{r}

fit_time_hos0 <- lm(time_in_hospital~.,data = diab_test)
summary(fit_time_hos0)

fit_time_hos1 <- lm(time_in_hospital~ race + num_medications + num_lab_procedures+ num_procedures + number_outpatient+
                  number_emergency + number_inpatient + number_diagnoses + A1Cresult + insulin + change + diabetesMed +
                  readmitted + new_age + new_adtype_id + new_discharge_id + new_adsource_id + new_med_spec + new_diag1+
                  new_diag2 + new_diag3, data = diab_test)
summary(fit_time_hos1)

fit_time_hos2 <- lm(time_in_hospital~ race + num_medications + num_lab_procedures+ num_procedures + number_outpatient+
                  number_emergency + number_inpatient + number_diagnoses + A1Cresult + insulin + diabetesMed +
                  readmitted + new_age + new_adtype_id + new_discharge_id + new_adsource_id + new_med_spec + new_diag1+
                  new_diag2 + new_diag3, data = diab_test)
summary(fit_time_hos2)
plot(fit_time_hos2,1)
plot(fit_time_hos2,5)

# The plot shows that the residuals are not randomly distributed, there are pattern I should work on to build a better model. Remove the variables which has correlation with num_medications.

fit_time_hos3 <- lm(time_in_hospital~ race + num_medications + number_outpatient+ number_emergency + 
                  number_inpatient + A1Cresult + insulin + diabetesMed +
                  readmitted + new_age + new_adtype_id + new_discharge_id + new_adsource_id + new_med_spec + new_diag1+
                  new_diag2 + new_diag3, data = diab_test)

summary(fit_time_hos3)
plot(fit_time_hos3,1)
plot(fit_time_hos3,5)

fit_time_hos4 <- lm(time_in_hospital~ race + num_medications + number_inpatient + number_emergency + A1Cresult + 
                  insulin + diabetesMed + readmitted + new_age + new_adtype_id + new_discharge_id + new_adsource_id +
                  new_med_spec + new_diag1 + new_diag2 + new_diag3, data = diab_test)
plot(fit_time_hos4,1)
plot(fit_time_hos4,5)

fit_time_hos5 <- lm(time_in_hospital~ race + num_medications  + number_inpatient + A1Cresult + 
                  insulin + diabetesMed + readmitted + new_age + new_adtype_id + new_discharge_id + new_adsource_id +
                  new_med_spec + new_diag1 + new_diag2 + new_diag3, data = diab_test)
summary(fit_time_hos5)
plot(fit_time_hos5,1)
plot(fit_time_hos5,5)

fit_time_hos6 <- lm(time_in_hospital~ race + num_medications  + A1Cresult + insulin + diabetesMed + 
                  readmitted + new_age + new_adtype_id + new_discharge_id + new_adsource_id +
                  new_med_spec + new_diag1 + new_diag2 + new_diag3, data = diab_test)

summary(fit_time_hos6)
plot(fit_time_hos6,1)
plot(fit_time_hos6,5)

```

10. Logistic regression model of "readmitted < 30"

a. Excluding "readmitted > 30"
```{r}
diab_glm_1 <- diab_test[diab_test$readmitted == "<30" | diab_test$readmitted == "NO",]
summary(diab_glm_1)
readmit_lev2 <- c("NO","<30")
diab_glm_1$readmitted <- factor(diab_glm_1$readmitted, levels = readmit_lev2)
levels(diab_glm_1$readmitted)
diab_glm2 <- na.omit(diab_glm_1)
summary(diab_glm2)

fit_all <- glm(readmitted~., family=binomial(), data = diab_glm2)
summary(fit_all)

fit_null <- glm(readmitted ~ 1, family = binomial(), data = diab_glm2)
summary(fit_null)

fit_step = step(fit_null, scope=list(lower=fit_null, upper=fit_all),direction="both")


summary(fit_step)

fit_final<- glm(readmitted ~ number_diagnoses + new_age + 
               time_in_hospital + diabetesMed + new_adtype_id + new_diag1 + 
               num_procedures + A1Cresult + insulin+ new_diag2 + 
               new_med_spec , family=binomial(), data = diab_glm2)

summary(fit_final)

logisticPseudoR2s(fit_final) 
pR2(fit_final)
exp(coef(fit_final))
vif(fit_final)
plot(fit_final,1)
plot(fit_final,5)
```


b. Including "readmitted > 30"

This model is worse fitted than the model above.
```{r}
diab_glm_2 <- diab_test
summary(diab_glm_2)
readmit_lev3 <- c("NO","<30")
diab_glm_2$readmitted[diab_glm_2$readmitted == ">30"] <- "NO"
diab_glm_2$readmitted <- factor(diab_glm_2$readmitted, levels = readmit_lev3)
summary(diab_glm_2$readmitted)
summary(diab_glm_2)


diab_glm3 <- na.omit(diab_glm_2)
summary(diab_glm3)

fit_all <- glm(readmitted~., family=binomial(), data = diab_glm3)
summary(fit_all)

fit_null <- glm(readmitted ~ 1, family = binomial(), data = diab_glm3)
summary(fit_null)

fit_step = step(fit_null, scope=list(lower=fit_null, upper=fit_all),direction="both")


summary(fit_step)

fit_final2<- glm(formula = readmitted ~ number_inpatient + 
    time_in_hospital + new_age + diabetesMed + number_diagnoses + 
    new_diag1 + new_med_spec + num_procedures + number_emergency + 
    new_diag2 + A1Cresult, family = binomial(), data = diab_glm3)

summary(fit_final2)

logisticPseudoR2s(fit_final2) 

pR2(fit_final2)
exp(coef(fit_final2))
vif(fit_final2)
plot(fit_final2,1)
plot(fit_final2,5)
```




